---
title: "Course Project 1"
output: html_document
keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Read Data

```{r read_data}
a <- read.csv("activity.csv")
a$date <- as.Date(a$date, "%Y-%m-%d")
```

## 2. Number of steps v/s Date

```{r steps}
c <- subset(a, !is.na(steps))
b <- with(c, tapply(steps, date, sum))
barplot(b, xlab = "Date", ylab = "Number of steps taken")
with(c, tapply(steps, date, mean))
with(c, tapply(steps, date, median))
```

## 3. Daily Activity Pattern

```{r daily activity}
d <-  with(c, tapply(steps, interval, sum))
e <- d/length(unique(c$date))
plot(unique(c$interval), e, type = "l", xlab = "Interval", ylab = "Number of steps taken")
max(e)
```

## 4. Imputing Missing Values

```{r missing values}
sum(is.na(a$steps))
```

The method used for imputing the missing values was:-
(a) Firstly we set the missing dates as the mean value for all those dates having known value and make a data frame containing date along with total number of steps taken on that date
(b) Then for all the rows containing missing steps we take average value from the above data frame.

```{r missing}
f <- a
g <- data.frame(date = names(b), steps = b)
h <- unique(a$date)
g$date <- as.Date(g$date, "%Y-%m-%d")
k <- h[!(h %in% g$date)]
m <- mean(g$steps)
l <- data.frame(date = k, steps = rep(m, length(k)))
n <- rbind.data.frame(g,l)
z <- n[order(n$date),]
for(i in 1:length(f$steps))
{
    if(is.na(f[i,"steps"]))
    {
        o <- f[i,"date"]
        num <- sum(f$date == o)
        p <- match(o, z$date)
        f[i, "steps"] = z[p, "steps"]/num
    }
}
sum(is.na(f$steps))
j <- with(f, tapply(steps, date, sum))
barplot(j, xlab = "Date", ylab = "Number of steps taken")
with(f, tapply(steps, date, mean))
with(f, tapply(steps, date, median))
```

Mean and median values do not differ from the previous estimate only new values have been entered.
The impact of imputing missing values is that we are able to see bar plot for more number of days.


## 5. Activity Patterns in weekdays and weekends

```{r weekdays, fig.height=10}
a$type <- weekdays(a$date)
for(q in 1:length(a$type))
{
    if(a[q,"type"] %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
    {
        a[q,"type"] = "weekday"
    }
    else
    {
        a[q,"type"] = "weekend"
    }
}
r <- subset(a, !is.na(steps))
u <- subset(r, type == "weekday")
v <- subset(r, type == "weekend")
w <- with(u, tapply(steps, interval, mean))
x <- with(v, tapply(steps, interval, mean))
par(mfrow=c(2,1))
plot(unique(u$interval),w, type = "l", main = "Weekday", xlab = "Interval", ylab = "Number of steps taken")
plot(unique(v$interval),x, type = "l", main = "Weekend", xlab = "Interval", ylab = "Number of steps taken")
```

